---
title: "Bios 735 Final Project_Vectorautogression(1)"
author: "Yun Lin"
date: "4/30/2022"
output: html_document
---
Goal: fit Vectorautogression(1) model with happiness score and its 6 factors for a selected county to predict its future happiness.
Load the package first.
```{r}
library(bios735.g3)
```
```{r}
install.packages("lmtest")
install.packages("dynlm")
install.packages("dyn")

library(dynlm)
library(lmtest)
library(dyn)
```


```{r}
# Read in dataset
happiness=read.csv('happiness_long.csv')
# select a specific country by choosing the country_id
ctr_id=1
happi=happiness[happiness$country_id == ctr_id, ]
```
Fit the VAR(1) model with the selected country. 
First, define the variables as time series (ts) object
```{r }
start_y=2015
end_y=2019

#score
score <- ts(happi$score,
            start = c(start_y), 
            end = c(end_y), 
            frequency = 1)


#GDP
GDPperCapita <- ts(happi$GDPperCapita,
                   start = c(start_y), 
                   end = c(end_y), 
                   frequency = 1)

#socialsupport
socialSupport<- ts(happi$socialSupport,
                   start = c(start_y), 
                   end = c(end_y), 
                   frequency = 1)

#healthyLifeExp
healthyLifeExp<- ts(happi$healthyLifeExp,
                    start = c(start_y), 
                    end = c(end_y), 
                    frequency = 1)

#freedom
freedom<- ts(happi$freedom,
             start = c(start_y), 
             end = c(end_y), 
             frequency = 1)

#generosity
generosity<- ts(happi$generosity,
                start = c(start_y), 
                end = c(end_y), 
                frequency = 1)

#corruptionPerception
corruptionPerception<- ts(happi$corruptionPerception,
                          start = c(start_y), 
                          end = c(end_y), 
                          frequency = 1)

```
VAR_1_model Equation_1
```{r}
# score_t = beta_0 + beta_1 * score_(t-1) + beta_2 * corruption_(t-1) + u_t
VAR_EQ1 <-
  dynlm(
    score ~ L(score, 1) + L(corruptionPerception, 1),
    start = c(start_y),
    end = c(end_y)
  )
summary(VAR_EQ1)
```

VAR_1_model Equation_2
```{r}
# corruption_t = beta_0 + beta_1 * score_(t-1) + beta_2 * corruption_(t-1) + u_t
VAR_EQ2 <-
  dynlm(
    corruptionPerception ~ L(score, 1) + L(corruptionPerception, 1) ,
    start = c(start_y),
    end = c(end_y)
  )
summary(VAR_EQ2)

```
the VAR(1) model failed because of insufficient amount of data.
we only have 5 data points (5 years) in each series, the model would work better in the future if data of more years are appended. 

